---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

[![Travis build status](https://travis-ci.org/egnha/dub.svg?branch=master)](https://travis-ci.org/egnha/dub)
[![Coverage status](https://codecov.io/gh/egnha/dub/branch/master/graph/badge.svg)](https://codecov.io/github/egnha/dub?branch=master)
![Status](https://img.shields.io/badge/status-WIP-blue.svg)

# dub

_dub_ is a tiny, single-purpose R package for _unpacking assignment_: it provides an operator `%<=%` that enables you to assign components of a list (or vector) to names, via pattern matching. Think of `<=` as a pictograph representing multiple `<-`'s.

The pattern matching syntax, which comes from [Haskell](https://en.wikibooks.org/wiki/Haskell/Pattern_matching), is concise and flexible. It mirrors the semantics of `list()`.

```{r}
library(dub)

(one : two : three) %<=% 1:3
one
two
three

(x) %<=% list(list("x"))
((y)) %<=% list(list("y"))
x
y

(u : (v : w)) %<=% list(1, list(2, 3:4))
u
v
w

# Use . to drop specific components, ... to drop greedily (the _ in Haskell)
(. : width : ... : species) %<=% iris
head(width)
head(species)
```

## Installation

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("egnha/dub")
```

## Desiderata

The main objective of dub is to cover the most common cases of unpacking assignment:

- assigning the top-level components of a list
- unpacking nested components

The pattern matching syntax should be economical and avoid conflict with R's established semantics. The code should be short and easy to grasp. There should be no external package dependencies.

### Implementation

dub overloads `` `:` `` as a symbol concatenator, in a strictly localized context. There is no semantic conflict with the usual `` `:` ``, which is an operator of an altogether different type (namely, a range operator of integers).

The brevity of the code reflects the simplicity of unpacking assignment: the leaves of the left-hand side of `%<=%` are matched, via `<-`, with the leaves on the right-hand side. The only subtlety is to handle greedy matching and make `` `:` `` right-associative.

## Prior art

Unpacking/multiple assignment appears in other languages, e.g., [Python](https://docs.python.org/3/tutorial/datastructures.html#tuples-and-sequences) and [Javascript](https://developer.mozilla.org/en-US/docs/Web/JavaScript/Reference/Operators/Destructuring_assignment). While R has no such feature, using a custom operator to do this has long been a folklore method. To my knowledge, the earliest implementation is by [Gabor Grothendieck](https://stat.ethz.ch/pipermail/r-help/2004-June/053343.html) (2004).

I recommend the [zeallot](https://github.com/nteetor/zeallot) package by [Nate Teetor](https://github.com/nteetor) as a more robust and full-featured option for unpacking assignment.

## License

MIT Copyright © 2018 [Eugene Ha](https://github.com/egnha)
